---
title: "Mixed Effect"
author: "Niklas Lystrup Poulsen"
date: "2025-05-13"
output: 
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## LOADING DATASET AND VARIABLES FOR USE
```{r mixed‐model‐all, message=FALSE, warning=FALSE}
# 1) load packages
library(readxl)
library(lme4)

# 2) read & prep data in one go
df <- read_excel("/Users/niklaspoulsen/Downloads/Master_Data_Thesis.xlsx") 



# 3) factor‐convert
df$productid_num <- factor(df$productid_num)
df$year          <- factor(df$year)
df$`Operation Code` <- factor(df$`Operation Code`)
#df$country       <- factor(df$country)

```


```{r}
```{r mixed_effects_tables, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
library(readxl)
library(lme4)
library(tibble)    
library(broom.mixed)
library(dplyr)
library(knitr)
library(lmerTest)
library(kableExtra)
library(plm)
library(broom)
library(purrr)
library(forcats)
library(knitr)
library(dplyr)
library(stringr)
library(tidyr)
library(lmerTest)  
library(lfe)
library(lme4)
library(tinytex)
library(ggplot2)



# make sure these are factors
#df$country         <- factor(df$country)
df$productid_num   <- factor(df$productid_num)
df$year            <- factor(df$year)     # if you want year‐dummies as fixed
df$country            <- factor(df$country)     # if you want year‐dummies as fixed


```


## FIRST FIVE ROWS OF THE DATASET - PRE TRANSFORMATION
```{r preview-truncated, echo=FALSE, results='asis'}

df      <- read_excel("book1.xlsx")
preview <- head(df, 5)

truncate_cell <- function(x, max_len = 8, mark = "...") {
  s <- as.character(x)
  ifelse(str_length(s) > max_len,
         paste0(str_sub(s, 1, max_len - str_length(mark)), mark),
         s)
}

preview_trunc <- preview %>%
  mutate(across(where(is.character), ~ truncate_cell(.x)))

preview_trunc %>%
  kbl(
    format   = "latex",
    booktabs = TRUE,
    linesep  = "",
    caption  = "First Five Rows of the Dataset - Pre Transformation (truncated)"
  ) %>%
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 8
  )

```


## FIRST FIVE ROWS OF DATASET - POST TRANSFORMATION
```{r preview‐kable, echo=FALSE, results='asis'}

df      <- read_excel("Master_Data_Thesis.xlsx")
preview <- head(df, 5)

truncate_cell <- function(x, max_len = 12, mark = "...") {
  s <- as.character(x)
  ifelse(nchar(s) > max_len,
         paste0(substr(s, 1, max_len - nchar(mark)), mark),
         s)
}
preview_trunc <- preview %>%
  mutate(across(where(is.character), ~ truncate_cell(.x)))

preview_trunc %>%
  kbl(
    format    = "latex",
    booktabs  = TRUE,
    linesep   = "",
    caption   = "First Five Rows of the Dataset - Post Transformation (truncated)"
  ) %>%
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 8
  )



```


## VISUALIZATION - EXPLORING PRODUCTS
```{r, echo=FALSE}
# 2. Products by Year
prod_by_year <- df %>%
  distinct(ProductID, year) %>%
  count(year, name = "num_products")

kable(
  prod_by_year,
  caption  = "Products by Year",
  booktabs = TRUE
)


# 3. Products by Country (All Years)
prod_by_country <- df %>%
  distinct(ProductID, country) %>%
  count(country, name = "num_products")

kable(
  prod_by_country,
  caption  = "Products by Country (All Years)",
  booktabs = TRUE
)

# 4. Products by Year & Country
prod_year_country <- df %>%
  distinct(ProductID, year, country) %>%
  count(year, country, name = "num_products") %>%
  pivot_wider(
    names_from  = country,
    values_from = num_products,
    values_fill = 0
  )

kable(
  prod_year_country,
  caption  = "Number of Products by Year and Country",
  booktabs = TRUE
)

```

## VISUALIZATION - NO. OF UNIQUE PRODUCTS PER OP CODE
```{r}
```{r products-per-opcode, results='asis', echo=FALSE}
library(dplyr)
library(knitr)

# 2) Count distinct products per operation code
code_counts <- df %>%
  group_by(`Operation Code`) %>%                 # backticks for the space
  summarise(num_products = n_distinct(productid_num)) %>%
  ungroup()

# 3) Print a nice table with caption
kable(
  code_counts,
  caption  = "Number of Unique Products per Operation Code",
  booktabs = TRUE
)
```

## VISUALIZATION - PATTERN OF PRODUCTS IN COUNTRY
```{r}
```{r country-patterns, message=FALSE, echo=FALSE, warning=FALSE, results='asis'}
# 1. Build the pattern table
pattern_tbl <- df %>% 
  distinct(productid_num, country, year) %>% 
  mutate(pres = 1) %>% 
  pivot_wider(
    names_from   = year,
    values_from  = pres,
    values_fill  = 0,
    names_prefix = "yr_"
  ) %>% 
  mutate(
    pattern = paste0(
      if_else(yr_2022 == 1, "1", "."),
      if_else(yr_2023 == 1, "1", "."),
      if_else(yr_2024 == 1, "1", ".")
    )
  ) %>% 
  count(country, pattern, name="freq") %>% 
  group_by(country) %>% 
  arrange(country, desc(pattern)) %>% 
  mutate(
    percent     = freq / sum(freq) * 100,
    cum_percent = cumsum(percent)
  ) %>% 
  ungroup()

# 2. For each country, cat header + kable
for(ctry in unique(pattern_tbl$country)) {
  tab <- pattern_tbl %>% 
    filter(country == ctry) %>% 
    select(pattern, freq, percent, cum_percent) %>% 
    rename(
      Pattern = pattern,
      Freq    = freq,
      Percent = percent,
      `Cum.`  = cum_percent
    )
  
  cat("\n\n### Country:", ctry, "\n\n")
  # the paste0 wrapper with collapse ensures knitr sees it properly
  cat(knitr::kable(tab, digits = c(0,0,2,2)), sep = "\n")
}
```




## SIMPLE REGRESSION + "U" VISUALIZATION
```{r plot‐residuals, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure Simple regression of $y$ on $x$ with residuals (error terms) $u_i$ shown as dashed lines."}
library(ggplot2)

# 1) Simulate example data
set.seed(2025)
df_simple <- data.frame(
  x = seq(-2, 2, length.out = 40)
)
df_simple$y <- 1.5 * df_simple$x + rnorm(nrow(df_simple), sd = 0.8)

# 2) Fit OLS
model <- lm(y ~ x, data = df_simple)
df_simple$yhat  <- predict(model)
df_simple$resid <- df_simple$y - df_simple$yhat

# 3) Plot points, fit line, and residuals
ggplot(df_simple, aes(x = x, y = y)) +
  geom_point(size = 2) +
  geom_abline(
    intercept = coef(model)[1],
    slope     = coef(model)[2],
    color     = "blue",
    size      = 1
  ) +
  geom_segment(
    aes(x = x, xend = x, y = y, yend = yhat),
    linetype = "dashed",
    color    = "red"
  ) +
  labs(
    x     = "Independent variable (x)",
    y     = "Dependent variable (y)",
    title = "Visualization of the Error Term u in Simple Regression"
  ) +
  theme_minimal()

```


## VISUALIZATION - No. of unique products per Operation Code
```{r}
```{r products-per-oppcode, results='asis', echo=FALSE}
library(dplyr)
library(knitr)

# 2) Count distinct products per operation code
code_counts <- df %>%
  group_by(`Operation Code`) %>%                 # backticks for the space
  summarise(num_products = n_distinct(productid_num)) %>%
  ungroup()

# 3) Print a nice table with caption
kable(
  code_counts,
  caption  = "Number of Unique Products per Operation Code",
  booktabs = TRUE
)
```
```

\newpage
## POOLED OLS PER COUNTRY
```{r}
```{r pooled‐ols‐by‐country, message=FALSE, echo=FALSE, warning=FALSE}
library(dplyr)
library(broom)
library(lmtest)
library(sandwich)
library(knitr)

# make sure operationcode is a factor


# 1) Run pooled OLS by country, clustering on productid_num
results <- df %>%
  group_by(country) %>%
  do({
    dat <- .

    # include operationcode dummies + year dummies
    mod <- lm(
      revenue ~ lag_miss + for_miss
              + factor(`Operation Code`)
              + factor(year),
      data = dat
    )

    # cluster‐robust SEs at the product level
    vc <- sandwich::vcovCL(mod,
                           cluster = dat$productid_num,
                           type = "HC0")   # match STATA's default robust

    # extract coefficient table
    ct <- lmtest::coeftest(mod, vcov = vc)
    broom::tidy(ct) %>%
      mutate(country = unique(dat$country))
  }) %>%
  ungroup()

# 2) Show the Denmark table
results %>%
  filter(country == "DK") %>%
  select(term, estimate, std.error, statistic, p.value) %>%
  kable(
    digits  = 3,
    caption = "Pooled OLS for Denmark (DK)"
  )

```

\newpage
```{r}
```{r pooled‐ols‐by‐countryy, message=FALSE, echo=FALSE, warning=FALSE}
library(dplyr)
library(broom)
library(lmtest)
library(sandwich)
library(knitr)

# make sure operationcode is a factor


# 1) Run pooled OLS by country, clustering on productid_num
results <- df %>%
  group_by(country) %>%
  do({
    dat <- .

    # include operationcode dummies + year dummies
    mod <- lm(
      revenue ~ lag_miss + for_miss
              + factor(`Operation Code`)
              + factor(year),
      data = dat
    )

    # cluster‐robust SEs at the product level
    vc <- sandwich::vcovCL(mod,
                           cluster = dat$productid_num,
                           type = "HC0")   # match STATA's default robust

    # extract coefficient table
    ct <- lmtest::coeftest(mod, vcov = vc)
    broom::tidy(ct) %>%
      mutate(country = unique(dat$country))
  }) %>%
  ungroup()

# 2) Show the Denmark table
results %>%
  filter(country == "SE") %>%
  select(term, estimate, std.error, statistic, p.value) %>%
  kable(
    digits  = 3,
    caption = "Pooled OLS for Sweden (SE)"
  )

```

\newpage
```{r}
```{r pooled‐ols‐by‐countries, message=FALSE, echo=FALSE, warning=FALSE}
library(dplyr)
library(broom)
library(lmtest)
library(sandwich)
library(knitr)

# make sure operationcode is a factor


# 1) Run pooled OLS by country, clustering on productid_num
results <- df %>%
  group_by(country) %>%
  do({
    dat <- .

    # include operationcode dummies + year dummies
    mod <- lm(
      revenue ~ lag_miss + for_miss
              + factor(`Operation Code`)
              + factor(year),
      data = dat
    )

    # cluster‐robust SEs at the product level
    vc <- sandwich::vcovCL(mod,
                           cluster = dat$productid_num,
                           type = "HC0")   # match STATA's default robust

    # extract coefficient table
    ct <- lmtest::coeftest(mod, vcov = vc)
    broom::tidy(ct) %>%
      mutate(country = unique(dat$country))
  }) %>%
  ungroup()

# 2) Show the Denmark table
results %>%
  filter(country == "DE") %>%
  select(term, estimate, std.error, statistic, p.value) %>%
  kable(
    digits  = 3,
    caption = "Pooled OLS for Germany (DE)"
  )

```

\newpage
```{r}
```{r pooled‐ols‐by‐countriess, message=FALSE, echo=FALSE, warning=FALSE}
library(dplyr)
library(broom)
library(lmtest)
library(sandwich)
library(knitr)

# make sure operationcode is a factor


# 1) Run pooled OLS by country, clustering on productid_num
results <- df %>%
  group_by(country) %>%
  do({
    dat <- .

    # include operationcode dummies + year dummies
    mod <- lm(
      revenue ~ lag_miss + for_miss
              + factor(`Operation Code`)
              + factor(year),
      data = dat
    )

    # cluster‐robust SEs at the product level
    vc <- sandwich::vcovCL(mod,
                           cluster = dat$productid_num,
                           type = "HC0")   # match STATA's default robust

    # extract coefficient table
    ct <- lmtest::coeftest(mod, vcov = vc)
    broom::tidy(ct) %>%
      mutate(country = unique(dat$country))
  }) %>%
  ungroup()

# 2) Show the Denmark table
results %>%
  filter(country == "NO") %>%
  select(term, estimate, std.error, statistic, p.value) %>%
  kable(
    digits  = 3,
    caption = "Pooled OLS for Norway (NO)"
  )



```


## FIXED-EFFECT
```{r}
```{r fe-tabless, echo=FALSE, results='asis'}
library(dplyr)
library(purrr)
library(plm)
library(sandwich)
library(lmtest)
library(broom)
library(kableExtra)

# 1. Estimate within-FE for each country and collect results
results_by_country <- df %>%
  split(.$country) %>%                     # list of dfs, one per country
  map_dfr(~ {
    fe_mod <- plm(
      revenue ~ lag_miss + for_miss + year,
      data   = .x,
      index  = c("productid_num","year"),
      model  = "within",
      effect = "individual"
    )
    vc_mat <- vcovHC(fe_mod, method="arellano", type="HC0", cluster="group")
    tidy(coeftest(fe_mod, vcov = vc_mat)) %>%
      mutate(country = unique(.x$country))
  }, .id = "country")

# 2. Print one LaTeX table per country
walk(unique(results_by_country$country), function(cty) {
  results_by_country %>%
    filter(country == cty) %>%
    select(term, estimate, std.error, statistic, p.value) %>%
    mutate(across(estimate:p.value, ~ round(.x, 3))) %>%
    kable(
      format    = "latex",
      booktabs  = TRUE,
      caption   = paste0("Fixed-Effect Estimates for ", cty, " (", cty, ")"),
      align     = c("l","r","r","r","r")
    ) %>%
    kable_styling(latex_options = "hold_position") %>%
    print()
})

```


